---
title: "Remote LWF-Brook90 run"
author: "Thomas Zieher"
date: "2025-06-04"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r header,include=T}
#####################################################
##                                                 ##
##      ######      #########  ##          ##      ##
##      #######     ########   ##          ##      ##
##      ##    ##    ##         ##          ##      ##
##      ##   ##     ##         ##          ##      ##
##      ######      ######     ##          ##      ##
##      #######     #####      ##          ##      ##
##      ##    ##    ##         ##    ##    ##      ##
##      ##     ##   ##         ##   ####   ##      ##
##      ##     ##   ##         ##  ##  ##  ##      ##
##      ########    ##          ####    ####       ##
##      #######     ##           ##      ##        ##
##                                                 ##
#####################################################
##                                                 ##
##  Author: Thomas Zieher                          ##
##  Institute: Nat Haz, 6.3                        ##
##  Date: 2025-05-28                               ##
##  License: GNU GPL V.3 or later                  ##
##                                                 ##
#####################################################
##  Description: Run LWF-Brook90 at San Vito di Cadore
##  Environment: renv
#####################################################

##set include=F to omit header in output
##code partly taken and modified after Schmidt-Walter et al. (2020), Supplement

```


```{r setup,include=FALSE}
knitr::opts_chunk$set(echo=T)

##empty environment
rm(list=ls())

##install package renv
#install.packages("renv")

##install packages with respective version
#renv::snapshot()
renv::restore()


library(LWFBrook90R)
library(data.table)
library(xts)
library(dygraphs)
library(ggplot2)
library(climatol)
library(rstudioapi)


##retrieving script path
proj_path=dirname(rstudioapi::getActiveProject())

##set paths
root=paste(proj_path,"earth_sensing_summer_school",sep="/")
model_path=paste(root,"data/lwf_brook90",sep="/")
meteo_path=paste(root,"data/era5",sep="/")
soil_path=paste(root,"data/soilgrids",sep="/")

##set period
start_date=as.Date("2019-01-01")
end_date=as.Date("2024-12-31")

```


***
### Read meteo data

```{r meteo,echo=T,message=F,warning=F,error=F}

##read ERA5 meteo data
meteo_daily=data.table(read.table(paste(meteo_path,"ERA5_meteo_daily.csv",sep="/"),header=T,sep=";",stringsAsFactors=F))
meteo_daily$dates=as.Date(meteo_daily$dates)

# ##adjust meteo
# meteo_daily$prec=meteo_daily$prec*0.5
# meteo_daily$tmean=meteo_daily$tmean+5.0
# meteo_daily$tmin=meteo_daily$tmin+5.0
# meteo_daily$tmax=meteo_daily$tmax+5.0


##Walther-Lieth diagram
##aggregate to monthly values
tmin=setDT(meteo_daily)[year(dates)%in%year(start_date):year(end_date), .(tmin=min(tmin)),by=.(yr=year(dates),mon=month(dates))]
tmax=setDT(meteo_daily)[year(dates)%in%year(start_date):year(end_date), .(tmax=max(tmax)),by=.(yr=year(dates),mon=month(dates))]
psum=setDT(meteo_daily)[year(dates)%in%year(start_date):year(end_date), .(psum=sum(prec)),by=.(yr=year(dates),mon=month(dates))]

##aggregate to mean monthly values
meteo_monthly=aggregate(data.frame(psum=psum$psum,
                                   tmax=tmax$tmax,
                                   tmin=tmin$tmin,
                                   tabsmin=abs(tmin$tmin)),
                        by=list(tmin$mon),FUN="mean")

par(mfrow=c(1,1))
diagwl(t(meteo_monthly[,2:5]), est="Location",alt="XXX",mlab="en",cols=NULL)


# hist(meteo_daily$prec)
# plot(meteo_daily$prec,type="h")
# aggregate(psum$psum,by=list(psum$yr),FUN="sum")
##too much precipitation, too cold
##grid point at high elevation??

```


***
### Parameterization
General options

```{r options,echo=T,message=F,warning=F,error=F}

##model options
options_forest=set_optionsLWFB90(startdate=start_date,
                                 enddate=end_date,
                                 budburst_method="Menzel",
                                 leaffall_method="vonWilpert",
                                 root_method="betamodel",
                                 lai_method="b90")

```


Soil profile

```{r soil,echo=T,message=F,warning=F,error=F}

##import soilgrids database
soilgrid_profiles=readRDS(paste(soil_path,"soilgrids_profiles.rds",sep="/"))
sites=aqp::site(soilgrid_profiles)
profiles=aqp::horizons(soilgrid_profiles)

##get single location
pid="P1"
location=sites[sites$id==pid,]
soil=profiles[profiles$id==pid,]

##modifications for LWF-Brook90
soil["upper"]=soil$upr/100.0*-1.0
soil["lower"]=soil$lwr/100.0*-1.0


# ##individual soil parameterization
# nlayers=6
# max_depth=-0.5#-0.2##-1##
# soil_raw=data.table()
# soil_raw$lower=max_depth
# soil_loc=soil_raw[rep(seq(1,nrow(soil_raw)),nlayers)]
# soil_loc$hor_num=1:nlayers
# soil_loc$gravel=0.8
# soil_loc$corg=2
# soil_loc$topsoil=rep(c(1,0),each=(nlayers/2))
# soil_loc$bd=1.0##1.5##2.0##
# soil_loc$upper=seq(0,(max_depth-(soil_raw$lower/nlayers)),length.out=nlayers)
# soil_loc$lower=seq((soil_raw$lower/nlayers),max_depth,length.out=nlayers)
# soil_loc$dw=15
# 
# soil_loc$clay=30
# soil_loc$silt=50
# soil_loc$sand=20
# 
# soil_loc=data.table(cbind(soil_loc,hydpar_hypres(clay=soil_loc$clay,
#                                                  silt=soil_loc$silt,
#                                                  bd=soil_loc$bd,
#                                                  oc.pct=soil_loc$corg,
#                                                  topsoil=soil_loc$topsoil)))
# 
# soil=soil_loc

```


Vegetation parameters

```{r vegetation,echo=T,message=F,warning=F,error=F}

##load parameterizations for European beech and Norway spruce from other project
load(paste(model_path,"veg_param.rda",sep="/"))

# ##individual vegetation parameterization
# ##taken from Schmidt-Walter et al. (2020), Supplement
# params_spruce=set_paramLWFB90(budburst_species="Picea abies (frueh)",
#                               maxlai=6,
#                               winlaifrac=0.8,
#                               height=25,
#                               sai=0.035*25,
#                               glmax=0.0053,
#                               radex=0.5,
#                               alb=0.14,
#                               albsn=0.14,
#                               maxrlen=3100,
#                               betaroot=0.976,
#                               maxrootdepth=-1.2,
#                               lwidth=0.004,
#                               frintlai=0.12,
#                               frintsai=0.14,
#                               cintrl=0.2,
#                               cintrs=0.4
#                               )
# 
# params_beech=set_paramLWFB90(budburst_species="Fagus sylvatica",
#                              maxlai=6,
#                              winlaifrac=0,
#                              height=25,
#                              sai=0.035*25,
#                              glmax=0.0053,
#                              radex=0.6,
#                              alb=0.18,
#                              albsn=0.23,
#                              maxrlen=3500,
#                              betaroot=0.966,
#                              maxrootdepth=-1.4,
#                              lwidth=0.05,
#                              frintlai=0.12,
#                              frintsai=0.25,
#                              cintrl=0.2,
#                              cintrs=0.4
#                              )

```


Site specifics

```{r site,echo=T,warning=F,error=F}

coords_x=12.215
coords_y=46.449
aspect=290 #east-north-east
slope=25 #deg
rooting_depth=-0.5 #m
drain=1 ##Switch for lower boundary condition to be free drainage (1) or no flow (0)
ilayer=4 ##Number of layers from top to which infiltration is distributed
infexp=0.6 ##Shape parameter for distribution of infiltration, for value 0 infiltration is in top layer only.

##set location's coordinates
params_spruce$coords_x=coords_x
params_spruce$coords_y=coords_y
params_beech$coords_x=coords_x
params_beech$coords_y=coords_y

##set location's aspect and slope angle
params_spruce$aspect=aspect
params_spruce$dslope=slope
params_spruce$eslope=slope
params_beech$aspect=aspect
params_beech$dslope=slope
params_beech$eslope=slope
  
##set location's rooting depth
params_spruce$maxrootdepth=rooting_depth
params_beech$maxrootdepth=rooting_depth

##set location's drainage parameter
params_spruce$drain=drain
params_beech$drain=drain

##set location's ilayer parameter
params_spruce$ilayer=ilayer
params_beech$ilayer=ilayer

##set location's infexp parameter
params_spruce$infexp=infexp
params_beech$infexp=infexp


```



***
### Run model

```{r model,echo=F,warning=F,error=F}

res_spruce=run_LWFB90(options_b90=options_forest,
                      param_b90=params_spruce,
                      climate=meteo_daily,
                      soil=soil
                      )

res_beech=run_LWFB90(options_b90=options_forest,
                     param_b90=params_beech,
                     climate=meteo_daily,
                     soil=soil
                     )

```


***
### Plots
Daily water balance components

```{r wb_daily,echo=F,message=F,warning=F,error=F}

##bind evap-fluxes from result lists
wb_spruce=rbindlist(list(`Spruce`=res_spruce$output),
                         idcol="vegetation")

wb_beech=rbindlist(list(`Beech`=res_beech$output),
                        idcol="vegetation")

##add date
wb_spruce$date=as.Date(sprintf("%i-%0.2i-%0.2i",wb_spruce$yr,wb_spruce$mo,wb_spruce$da),format="%Y-%m-%d")
wb_beech$date=as.Date(sprintf("%i-%0.2i-%0.2i",wb_beech$yr,wb_beech$mo,wb_beech$da),format="%Y-%m-%d")

##add surface runoff
wb_spruce[,surfrunoff := flow-vrfln]
wb_beech[,surfrunoff := flow-vrfln]

##bind results
wb=rbind(wb_spruce,wb_beech)

##rfal and smlt from model results
tmin=xts(meteo_daily$tmin,order.by=meteo_daily$dates)
tmax=xts(meteo_daily$tmax,order.by=meteo_daily$dates)
tmean=xts(meteo_daily$tmean,order.by=meteo_daily$dates)

nprec=xts(wb_beech$rfal,order.by=wb_beech$date)
sfal=xts(wb_beech$sfal,order.by=wb_beech$date)
smlt=xts(-wb_beech$smlt,order.by=wb_beech$date)
srf=xts(-wb_beech$surfrunoff,order.by=wb_beech$date)

tseries_plot=merge(tmin,tmax,tmean,nprec,sfal,smlt,srf)

##fill missing 15min measurements with closest measurement
#tseries_smc_filled=na.locf(tseries_plot,fromLast=F,maxgap=10)

##plot wb
cols_prec=c("#0800ff","#50aaff","#50ffff","#F39200","#ebeb34","#F39200","#FF1000")
par(mar=c(5,5,5,5))
dygraph(tseries_plot,width=600,height=400) %>%
  dyAxis("y",label="Water balance elements [mm]",valueRange=c(-20,60)) %>%
  dyAxis("y2",label="Temperature [deg C]",valueRange=c(-10,30)) %>%
  dyBarSeries('nprec',label="Rainfall [mm/d]",axis="y") %>% 
  dyBarSeries('sfal',label="Snow fall [mm/d]",axis="y") %>% 
  dyBarSeries('smlt',label="Snow melt [mm/d]",axis="y") %>% 
  dyBarSeries('srf',label="Surface runoff [mm/d]",axis="y") %>% 
  dyOptions(stackedGraph=T) %>%
  dySeries('tmin',label="Tmin [deg C]",axis="y2") %>% 
  dySeries('tmean',label="Tmean [deg C]",axis="y2") %>% 
  dySeries('tmax',label="Tmax [deg C]",axis="y2") %>% 
  dyOptions(stackedGraph=F,colors=cols_prec) %>%
  dyRangeSelector(height=20) %>%
  dyLegend(show="always",width=120)


```


```{r swati,echo=F,warning=F,error=F}

##bind evap-fluxes from result lists
swat_spruce=rbindlist(list(`Spruce`=res_spruce$layer_output),
                      idcol="vegetation")

swat_beech=rbindlist(list(`Beech`=res_beech$layer_output),
                     idcol="vegetation")

```


***
### T_ratio per DOY
Values <Q01 are shown as point markers

```{r tdiff_plot,echo=F,warning=F,error=F,out.width="80%",fig.width=10,fig.height=13}

##re-read outputs
output_beech=wb_beech
output_spruce=wb_spruce

col_beech="#ff916e"
col_spruce="#159837"
col_mixed="#a387ff"
cols_tratio=c(col_beech,col_spruce,col_mixed)

##BEECH
output_beech$tratio=output_beech$tran/output_beech$ptran
output_beech$tratio[output_beech$tratio>1]=1
output_beech$tratio[is.nan(output_beech$tratio)]=1
output_beech$tratio[is.infinite(output_beech$tratio)]=1

##SPRUCE
output_spruce$tratio=output_spruce$tran/output_spruce$ptran
output_spruce$tratio[output_spruce$tratio>1]=1
output_spruce$tratio[is.nan(output_spruce$tratio)]=1
output_spruce$tratio[is.infinite(output_spruce$tratio)]=1


##mixed forest (mean of beech and spruce)
mixed=data.frame(doy=output_beech$doy,
                 yr=output_beech$yr,
                 date=output_beech$date,
                 rnet=apply(cbind(output_beech$rnet,output_spruce$rnet),1,FUN="mean",na.rm=T),
                 smlt=apply(cbind(output_beech$smlt,output_spruce$smlt),1,FUN="mean",na.rm=T),
                 tratio=apply(cbind(output_beech$tratio,output_spruce$tratio),1,FUN="mean",na.rm=T)
)


##plot all years
par(mfrow=c(6,1),mar=c(2,4,3,1),pty="m")
#year_plot=unique(output_beech$yr)[1]
for (year_plot in unique(output_beech$yr)){
  beech_sub=output_beech[output_beech$yr==year_plot,]
  spruce_sub=output_spruce[output_spruce$yr==year_plot,]
  mixed_sub=mixed[mixed$yr==year_plot,]
  
  plot(NA,NA,xlim=c(1,366),ylim=c(0,1),xlab="",ylab="Transpiration ratio",main=sprintf("Transpiration ratio %i",year_plot))
  
  lines(beech_sub$doy,beech_sub$tratio,col=col_beech)
  lines(spruce_sub$doy,spruce_sub$tratio,col=col_spruce)
  lines(mixed_sub$doy,mixed_sub$tratio,col=col_mixed)

  points(beech_sub$doy[beech_sub$tratio<quantile(na.omit(beech_sub$tratio),0.01)],
         beech_sub$tratio[beech_sub$tratio<quantile(na.omit(beech_sub$tratio),0.01)],col=col_beech,pch=1)
  points(spruce_sub$doy[spruce_sub$tratio<quantile(na.omit(spruce_sub$tratio),0.01)],
         spruce_sub$tratio[spruce_sub$tratio<quantile(na.omit(spruce_sub$tratio),0.01)],col=col_spruce,pch=3)
  points(mixed_sub$doy[mixed_sub$tratio<quantile(na.omit(mixed_sub$tratio),0.01)],
         mixed_sub$tratio[mixed_sub$tratio<quantile(na.omit(mixed_sub$tratio),0.01)],col=col_mixed,pch=5)
  
  
  ##plot thresholds
  abline(h=0.25,col="red",lty=2)
  abline(h=0.75,col="orange",lty=2)
  
  legend("bottomright",legend=c("Beech","Spruce","Mixed"),lwd=2,pch=1:5,col=cols_tratio)
  
}


##dygraph
##rfal and smlt from model results
Beech=xts(output_beech$tratio[output_beech$yr%in%unique(output_beech$yr)],order.by=output_beech$date[output_beech$yr%in%unique(output_beech$yr)])
Spruce=xts(output_spruce$tratio[output_spruce$yr%in%unique(output_beech$yr)],order.by=output_spruce$date[output_spruce$yr%in%unique(output_beech$yr)])
Mixed=xts(mixed$tratio[mixed$yr%in%unique(output_beech$yr)],order.by=mixed$date[mixed$yr%in%unique(output_beech$yr)])

prec=xts(mixed$rnet[mixed$yr%in%unique(output_beech$yr)],order.by=mixed$date[mixed$yr%in%unique(output_beech$yr)])
smlt=xts(mixed$smlt[mixed$yr%in%unique(output_beech$yr)],order.by=mixed$date[mixed$yr%in%unique(output_beech$yr)])

tseries_plot=merge(prec,smlt,Beech,Spruce,Mixed)

##fill missing 15min measurements with closest measurement
#tseries_smc_filled=na.locf(tseries_plot,fromLast=F,maxgap=10)

##plot
cols=c("#0800ff","#44bbff",cols_tratio)
par(mar=c(5,5,5,5))
dygraphs::dygraph(tseries_plot,width=600,height=400) %>%
  dyAxis("y",label="Transpiration ratio",valueRange=c(0,1)) %>%
  dyAxis("y2",label="Net rainfall & snow melt (mixed forest) [mm]",valueRange=c(50,0)) %>%
  dyGroup(c("Beech","Spruce","Mixed"),drawPoints=F,color=cols_tratio,axis="y") %>%
  dyLimit(0.25,strokePattern="dashed",color="red") %>%
  dyLimit(0.75,strokePattern="dashed",color="orange") %>%
  dyOptions(stackedGraph=F,colors=cols) %>%
  dyBarSeries('prec',label="Net rainfall [mm/d]",axis="y2") %>% 
  dyBarSeries('smlt',label="Snow melt [mm/d]",axis="y2") %>%
  dyRangeSelector(height=20) %>%
  dyLegend(show="always",width=120)



```


***

### Relative extractable water (REW) per DOY
Values <Q01 are shown as point markers

```{r rew_plot,echo=F,warning=F,error=F,out.width="80%",fig.width=10,fig.height=13}

mixed_rew=data.frame(doy=output_beech$doy,
                     yr=output_beech$yr,
                     date=output_beech$date,
                     rnet=apply(cbind(output_beech$rnet,output_spruce$rnet),1,FUN="mean",na.rm=T),
                     smlt=apply(cbind(output_beech$smlt,output_spruce$smlt),1,FUN="mean",na.rm=T),
                     relawat=apply(cbind(output_beech$relawat,output_spruce$relawat),1,FUN="mean",na.rm=T)
)

##plot all years
par(mfrow=c(6,1),mar=c(2,4,3,1),pty="m")
#year_plot=unique(output_beech$yr)[1]
for (year_plot in unique(output_beech$yr)){
  beech_sub=output_beech[output_beech$yr==year_plot,]
  spruce_sub=output_spruce[output_spruce$yr==year_plot,]
  mixed_sub=mixed_rew[mixed_rew$yr==year_plot,]
  
  plot(NA,NA,xlim=c(1,366),ylim=c(0,2),xlab="",ylab="Relative extractable water",main=sprintf("Relative extractable water %i",year_plot))
  
  lines(beech_sub$doy,beech_sub$relawat,col=col_beech)
  lines(spruce_sub$doy,spruce_sub$relawat,col=col_spruce)
  lines(mixed_sub$doy,mixed_sub$relawat,col=col_mixed)

  points(beech_sub$doy[beech_sub$relawat<quantile(na.omit(beech_sub$relawat),0.01)],
         beech_sub$relawat[beech_sub$relawat<quantile(na.omit(beech_sub$relawat),0.01)],col=col_beech,pch=1)
  points(spruce_sub$doy[spruce_sub$relawat<quantile(na.omit(spruce_sub$relawat),0.01)],
         spruce_sub$relawat[spruce_sub$relawat<quantile(na.omit(spruce_sub$relawat),0.01)],col=col_spruce,pch=3)
  points(mixed_sub$doy[mixed_sub$relawat<quantile(na.omit(mixed_sub$relawat),0.01)],
         mixed_sub$relawat[mixed_sub$relawat<quantile(na.omit(mixed_sub$relawat),0.01)],col=col_mixed,pch=5)
  
  
  ##plot thresholds
  abline(h=0.2,col="red",lty=2)
  abline(h=0.4,col="orange",lty=2)
  abline(h=1,col="green",lty=2)
  
  legend("topright",legend=c("Beech","Spruce","Mixed"),lwd=2,pch=1:5,col=cols_tratio)
  
}



##dygraph
##rfal and smlt from model results
Beech=xts(output_beech$relawat[output_beech$yr%in%unique(output_beech$yr)],order.by=output_beech$date[output_beech$yr%in%unique(output_beech$yr)])
Spruce=xts(output_spruce$relawat[output_spruce$yr%in%unique(output_beech$yr)],order.by=output_spruce$date[output_spruce$yr%in%unique(output_beech$yr)])
Mixed=xts(mixed_rew$relawat[mixed_rew$yr%in%unique(output_beech$yr)],order.by=mixed_rew$date[mixed_rew$yr%in%unique(output_beech$yr)])

prec=xts(mixed_rew$rnet[mixed_rew$yr%in%unique(output_beech$yr)],order.by=mixed_rew$date[mixed_rew$yr%in%unique(output_beech$yr)])
smlt=xts(mixed_rew$smlt[mixed_rew$yr%in%unique(output_beech$yr)],order.by=mixed_rew$date[mixed_rew$yr%in%unique(output_beech$yr)])

tseries_plot=merge(prec,smlt,Beech,Spruce,Mixed)

##fill missing 15min measurements with closest measurement
#tseries_smc_filled=na.locf(tseries_plot,fromLast=F,maxgap=10)

dygraphs::dygraph(tseries_plot,width=600,height=400) %>%
  dyAxis("y",label="Relative extractable water (REW)",valueRange=c(0,2)) %>%
  dyAxis("y2",label="Net rainfall & snow melt (mixed forest) [mm]",valueRange=c(50,0)) %>%
  dyGroup(c("Beech","Spruce","Mixed"),drawPoints=F,color=cols_tratio,axis="y") %>%
  dyLimit(0.2,strokePattern="dashed",color="red") %>%
  dyLimit(0.4,strokePattern="dashed",color="orange") %>%
  dyLimit(1,strokePattern="dashed",color="green") %>%
  dyOptions(stackedGraph=F,colors=cols) %>%
  dyBarSeries('prec',label="Net rainfall [mm/d]",axis="y2") %>% 
  dyBarSeries('smlt',label="Snow melt [mm/d]",axis="y2") %>%
  dyRangeSelector(height=20) %>%
  dyLegend(show="always",width=120)


```


***
### Water balance: Mean monthly conditions

```{r annual,echo=F,message=F,warning=F,error=F}

##compute monthly sums and means
##"tdiff","evap","tran", "ievp", "gevp","surfrunoff","vrfln", "flow"
mean_list=c("snow","swat","gwat","seep","gwfl","srfl","slfl","byfl","dsfl",
                  "stres","adef","awat","relawat","nits","balerr","slrad","solnet","lngnet","aa","asubs")

sum_list=c("rfal","rint","sfal","sint","rthr","sthr","rsno","rnet",
           "smlt","intr","ints","evap","tran",
           "irvp","isvp","slvp","snvp",
           "pint","ptran","pslvp","safrac",
           "surfrunoff","flow","vrfln"
           )

##aggregate values to sum and mean
agg_sum=aggregate(wb[,..sum_list],by=list(sprintf("%i-%0.2i-01",wb$yr,wb$mo),
                                               wb$vegetation),FUN="sum")
agg_sum["date"]=as.Date(agg_sum[,1],format="%Y-%m-%d")
agg_sum["yr"]=year(agg_sum$date)
agg_sum["mo"]=month(agg_sum$date)
agg_sum["vegetation"]=agg_sum[,2]
agg_sum=agg_sum[!grepl(pattern="Group",names(agg_sum))]

agg_mean=aggregate(wb[,..mean_list],by=list(sprintf("%i-%0.2i-01",wb$yr,wb$mo),
                                                 wb$vegetation),FUN="mean")
agg_mean=agg_mean[!grepl(pattern="Group",names(agg_mean))]

##bind monthly aggregated values
evapmon=data.table(cbind(agg_sum,agg_mean))

##aggregate all variables to monthly values
evap_m=evapmon[yr %in% year(start_date):year(end_date),lapply(.SD, mean),by=list(vegetation,mo)]

##melt
hgts=melt(evap_m[,list(vegetation,mo,
                          evap=evap,
                          gevp=slvp+snvp,
                          ievp=irvp+isvp,
                          tran,
                          tdiff=ptran-tran,
                          vrfln=-vrfln,
                          surfrunoff=-surfrunoff,
                          flow=-flow)],
          id.vars=c("vegetation","mo"))


hgts$month=factor(month.abb[hgts$mo],levels=month.abb)
hgts$variable=factor(hgts$variable,levels=c("tdiff","evap","tran", "ievp", "gevp","surfrunoff","vrfln","flow"))

##precipitation
prec=meteo_daily[year(dates) %in% year(start_date):year(end_date),list(prec=sum(prec)),
           by=list(year(dates),month(dates))][,lapply(.SD,mean),by=month]
prec$month=factor(month.abb[prec$month],levels=month.abb)


##bar colors
barcolors=c("#d01c8b","#abdda4","#ffffbf","#fdae61","#91bfdb","#2b83EE","#2b83cf","#2b83ba")

##theme for plotting
theme_budg=theme(
  axis.text.x=element_text(angle=90,vjust=0.5),
  axis.line=element_line(linewidth=0.3),
  axis.ticks=element_line(linewidth=0.3),
  legend.position="right",
  legend.background=element_blank(),
  legend.key=element_blank(),
  panel.background=element_rect(color="black",fill=NA),
  panel.grid=element_blank(),
  panel.border=element_rect(colour="black",fill=NA,size=0.3),
  strip.background.x=element_rect(color="black",fill=NA,size=0.3),
  axis.text=element_text(size=7),
  axis.title=element_text(size=10),
  legend.text=element_text(size=8),
  legend.key.size=unit(5,"mm"))

p_budg=ggplot(hgts[!variable %in% c("evap","flow"),]) +
  geom_bar(aes(x=month,y=value,fill=variable),
           stat='identity',position='stack',color="black",linewidth=0.05) +
  geom_line(data=prec,
            aes(x=month,y=prec,group=1,color="P"),linewidth=0.4) +
  scale_color_manual(values="black",labels="P",name="") +
  scale_fill_manual(values=barcolors,labels=expression(T[D],T[a],E[i],E[s],R,D)) +
  facet_grid(~ vegetation) +
  labs(x ="Month",y="Water flux [mm]",fill="",title=NULL) +
  scale_y_continuous(sec.axis=sec_axis(transform=~.*1),breaks=seq(-100,200,50))+
  theme_budg
print(p_budg)

```


